\documentclass[10pt, openright]{book}

% --- KDP 5x8 PAPERBACK CONFIGURATION ---
% 5x8 inch trim size.
% 'inner' margin is larger (0.75in) to account for the spine binding (gutter).
% 'outer' margin is smaller (0.5in) to maximize reading space.
\usepackage[paperwidth=5in, paperheight=8in, top=0.75in, bottom=0.75in, outer=0.5in, inner=0.75in]{geometry}

% --- TYPOGRAPHY ---
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{ebgaramond} % The Classic Serif Font (Time-honored and legible)
\usepackage{microtype}  % Essential for professional typesetting (smooths edges)
\usepackage{chngcntr}   % For continuous numbering

% --- STYLING PACKAGES ---
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{parskip}
\usepackage{amssymb}
\usepackage{emptypage} % Prevents page numbers/headers on blank pages

% --- 1. CHAPTER STYLING (The "Books") ---
	\titleformat{\chapter}[display]
  {\normalfont\centering} % Global style for chapter
  {\vspace{-1cm}\itshape\large Chapter \thechapter} % Label "Chapter I"
  {10pt} % Space between label and title
  {\Huge\bfseries} % Title style
	\titlespacing*{\chapter}{0pt}{0pt}{40pt}

% --- 2. SAYING STYLING (The "Statuary" Look) ---
% We use [display] to stack the number above the title
\counterwithout{section}{chapter} % Continuous numbering 1-24
	\titleformat{\section}[display]
  {\normalfont\centering} % Center everything
  {\vspace{1em}\Huge\bfseries\thesection} % THE NUMBER: Huge and Bold
  {0.3em} % Small gap between number and title
  {\large\scshape} % THE TITLE: Small Caps, slightly smaller than number
  
% Adjust spacing: {left}{top}{bottom}
	\titlespacing*{\section}{0pt}{1.0ex plus 1ex minus .2ex}{1.5ex plus .2ex}

% --- 3. PAGE BREAK LOGIC ---
% CRITICAL: This command ensures a saying is never split across pages.
% If it doesn't fit, it moves to the next page entirely.
\newcommand{\sectionbreak}{\filbreak}

% --- 4. ORNAMENTS ---
\newcommand{\ornament}{
    \vspace{1.5cm}
    \begin{center}
        \large $\cdot$ $\odot$ $\cdot$
    \end{center}
    \vspace{1.5cm}
}

% --- 5. HEADER/FOOTER ---
\pagestyle{fancy}
\fancyhf{}
\fancyhead[CO]{\scshape\small The Book of Awakening} 
\fancyhead[CE]{\scshape\small Yeshua the Living One}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0pt}

\begin{document}

\begin{titlepage}
    \centering
    \vspace*{1.5cm}
    {\Huge \textsc{The Book of}\\ \textsc{Awakening}} \\
    \vspace{0.5cm}
  {\large \textit{Suttas of the Living Buddha}} \\
    \vspace{2cm}
	    \textbf{The Path from Forgetting to Clarity} \\
    \vspace{0.5cm}
    {\small Edition 2.0 $\cdot$ December 5, 2025} \\
    \vfill
    {\small Transcribed for the One Who Seeks to See}
    \vspace{1cm}
\end{titlepage}

\frontmatter
\tableofcontents

\chapter{Prologue}
\textit{To the one who seeks to see\dots}

Yeshua said:

``First there is forgetting: the soul sleeps in the garment of the world.
Then there is awakening: the soul opens its eyes and sees that it is not the garment.
Finally there is remembering: the soul knows it is the Light itself.

This is the scroll of Awakening. Here we accept the duality of the watcher and the watched, not to remain divided, but to gain clarity. For you cannot be what you observe. You are the Seer, not the seen.''
Whenever you are struck, \\
ask: `What is the first arrow I cannot avoid, \\
and what is the second arrow I am adding?' \\
Release the second, \\
and offer the first to the Father in trust.\\

\ornament

\mainmatter

% =========================================================
% CHAPTER 1
% =========================================================
\chapter{The Breath of the Living Source}

\section{The Sacred Breath}
Yeshua said: \\
``The breath you call yours \\
was given before you had a name. \\
Follow it gently \\
back to its hidden spring, \\
and you will find the One \\
who breathes all things into being \\
and gathers them home again.''\\

\section{The Guest House}
If a thought grips you, \\
pause before you act. \\
Feel your feet upon the earth, \\
take three quiet breaths, \\
and let the river pass before you \\
like clouds across the sky.\\
``A thought knocks at your door \\
and pretends to be the master. \\
Do not drive it away; \\
do not fall at its feet. \\
Simply open the door of your seeing \\
and watch it come and go. \\
When you remember a hurt, \\
place your hand over your heart \\
and say, `Here is pain; may it be healed.' \\
Say the same for the one who wounded you. \\
In this, you step out of the chain \\
and stand in the Light.\\
The true Master \\
does not arrive or depart— \\
He was the space of the house \\
all along.''\\

\section{The Rhythm of the All}
Yeshua said: \\
``Every inhale is the world entering you. \\
Every exhale is your soul returning to the world. \\
Between these two tides \\
lies a stillness \\
where ocean and drop remember \\
they are one water. \\
Bring your whole attention \\
to this simple breath. \\
In that seeing, \\
the current of thoughts is cut \\
like a rope beneath a bell, \\
Choose one small, ordinary act each day— \\
washing a cup, tying a shoe— \\
and do it as if nothing else exists. \\
Let this be your gate into the Kingdom.\\
and only the clear ringing \\
of the present remains.''

\ornament

% =========================================================
% CHAPTER 2
% =========================================================
\chapter{Mindfulness of the Inner Light}

\section{The Lamp of the Mind}
Yeshua said: \\
``The mind is a lamp \\
set upon the altar of the heart. \\
When it turns toward many images, \\
shadows crowd the room. \\
Turn the lamp toward this present breath, \\
this single step, \\
Whenever fear of yesterday or tomorrow arises, \\
gently return to what your hands are touching now. \\
Let the present moment \\
be the ark that carries you through the waters.\\
this bare moment, \\
and its light becomes a sword— \\
cutting through drifting thoughts \\
as a sharp wind scatters smoke. \\
What survives the cut \\
is what is truly here.''\\

\section{The Seeing Itself}
``You see the sky, \\
so you are not the sky. \\
You see the thought as it arises now, \\
so you are not the thought. \\
You see even the sense of `I' \\
appearing and fading in this moment, \\
so you are not that either. \\
Rest your attention in the seeing itself— \\
still, awake, present— \\
and the Seer opens \\
like a flower of light \\
with no one at the center.''\\

\section{The Birth of Light}
``Sit until the silence \\
no longer seems a stranger. \\
Sit until the watcher grows thin \\
as morning mist. \\
Then you will know \\
that the Light you sought \\
was waiting, patient and untroubled, \\
beneath the next unspoken breath.''

\ornament

% =========================================================
% CHAPTER 3
% =========================================================
\chapter{The Nature of Suffering}

\section{The Shadow and the Light}
``A shadow follows your feet \\
and whispers, `I am you.' \\
But you are the sun \\
that casts the shadow— \\
untouched, \\
unbroken, \\
unbound. \\
Suffering begins \\
when you bow to the shadow. \\
Freedom begins \\
when you remember the Light.''\\

\section{The Garment of Desire}
``When desire arises, \\
let it rise like mist on a warm morning. \\
Feel it on the skin, \\
but do not call it by your name. \\
Touch it lightly, \\
as one touches a garment hanging loosely. \\
Remember: \\
you are the wearer, \\
not the woven thread. \\
Desire that is seen clearly \\
fades by itself \\
and leaves the heart clean.''\\

\section{Facing the Shadow}
``The first arrow is pain \\
when life pierces the body or the heart. \\
The second arrow \\
is the story you tell about it— \\
`This should not be,' \\
`This is who I am.' \\
You cannot always turn away \\
the first arrow. \\
But the second \\
is in your own hand. \\
Bring your attention gently \\
to the ache that is here now, \\
without name, \\
without judgment. \\
In that clear seeing, \\
the second arrow falls, \\
and the wound becomes a doorway. \\
Bow to the darkness with open eyes— \\
and it scatters \\
like crows before dawn.''

\ornament

% =========================================================
% CHAPTER 4
% =========================================================
\chapter{The Path of Seeing Through Illusion}

\section{The Ocean and the Wave}
Yeshua said: \\
``A thought rises— \\
a single wave \\
tasting the open air. \\
Chase it, \\
and you forget the sea. \\
Be still, \\
and the wave returns \\
to the depth \\
that has always been yours. \\
You are not the passing wave, \\
but the ocean that never leaves itself.''\\

\section{The Space Around the Fire}
``Anger flames high \\
when fed by the self. \\
Stand one breath away, \\
as the clear night \\
stands from the campfire. \\
See the heat, \\
feel the pull, \\
but do not step inside the blaze. \\
The fire burns itself out, \\
and the night \\
remains unburned. \\
So too your awareness \\
remains untouched \\
when you do not call the fire `I'.''\\

\section{The River of Thoughts}
Yeshua said: \\
``Thoughts flow through the mind \\
like a river in flood— \\
memories, fears, plans, \\
arguing for a thousand futures. \\
If you are swept into the current, \\
you are tossed from stone to stone. \\
If you stand on the bank \\
in the clear strength of the present, \\
you feel the rush, \\
but you do not drown. \\
Let your attention be a sword \\
of quiet seeing. \\
With a single, gentle stroke \\
it parts the waters of distraction, \\
and the bed of the river appears— \\
the still, unmoving ground \\
of your own being.''

\ornament

% =========================================================
% CHAPTER 5
% =========================================================
\chapter{Compassion Born of Non-Self}

\section{The Seeing of Compassion}
``Compassion is not the trembling \\
of a tender heart alone. \\
It is the clear sight \\
that no leaf \\
falls apart from the tree, \\
and no soul \\
stands apart from the One. \\
When you see through the veil of `me' and `you', \\
you cannot help but care— \\
for every face \\
is your own.''\\

\section{Healing the Inner Wound}
``When another wounds you, \\
listen for the wound \\
beneath their hand. \\
It is their cry, \\
not your identity. \\
Heal the echo within yourself— \\
the place that believed \\
it could be diminished— \\
and the blow transforms \\
into a bell \\
calling you home. \\
Then you will bless \\
even the one who struck you, \\
for unknowingly \\
they pointed to your freedom.''\\

\section{The Natural Radiance}
``When the self grows quiet, \\
love spills out \\
as naturally \\
as water reflecting the moon. \\
You need not command it, \\
nor adorn it with names. \\
You need only \\
step out of the way. \\
Then compassion moves by itself— \\
no doer, \\
only the shining of the Heart.''

\section{The Space of Seeing}
Yeshua said: \\
``When you look directly \\
at a thought, a fear, a wound, \\
it steps out from your chest \\
and stands before you. \\
In that simple seeing, \\
space opens. \\
What was a prison \\
becomes a guest you may receive \\
and gently release. \\
What you behold, \\
you are no longer bound to be.''\\

\section{The Four Lettings-Go}
``There are four doors \\
through which the false self leaves quietly: \\
non-violence, \\
non-judgment, \\
non-forcing, \\
and non-self. \\
When you do not wound, \\
the heart need not defend. \\
When you do not judge, \\
the mind need not divide. \\
When you do not force, \\
the Way moves of itself. \\
When you do not cling to `I', \\
the true "I Am" may speak. \\
You do not disappear— \\
you become clear.''\\

\section{The Union in the Heart}
``In the heart that has become clear \\
through these lettings-go, \\
the Father and the Son are not two. \\
The Source and the Living Image \\
shine as one Light. \\
The Holy Breath \\
moves between them \\
like fire between coal and flame, \\
and completes their joy in you. \\
This is the bridal chamber within: \\
when you cease to wound, to judge, \\
to force, to cling, \\
you do not become empty— \\
you become the open room \\
where God and the soul \\
know themselves as one.""

\ornament

% =========================================================
% CHAPTER 6
% =========================================================
\chapter{The Kingdom as the Present Moment}

\section{The Kingdom is Now}
Yeshua said: \\
``The Kingdom hides \\
in the folds of the present moment \\
like a jewel in a child's hand. \\
Do not search in distant heavens \\
or in years yet to come. \\
Bring your whole attention \\
to this breath, \\
this sound, \\
this simple act of standing here. \\
Such attention is a bright blade \\
that cuts the chains of yesterday and tomorrow. \\
Whoever stands thus in now \\
finds that the doors of the Kingdom \\
were never shut.''\\

\section{The Living Moment}
``The past is a footprint in dust. \\
The future is a story in ink, \\
yet to be read. \\
This breath alone is alive, \\
and the Living One \\
whispers within it. \\
Drink this breath without distraction, \\
and you will taste \\
the wine of eternity \\
poured into now.''\\

\section{The Dream of Time}
``When you step fully into now, \\
the watchers of fear fall silent. \\
For they live only \\
in yesterday's shadows \\
and tomorrow's storms. \\
Hold the present moment \\
as you would hold a flame— \\
steadily, with care. \\
Watched in this way, \\
the dream of time unravels \\
thread by thread, \\
until only naked Being remains. \\
Then you will see \\
that you never left eternity— \\
you only wandered in your thoughts.''

\ornament

% =========================================================
% CHAPTER: THE FRUIT OF THE WAY
% (On the fruit of action / fruit of being.)

\chapter{The Fruit of the Way}

\section{The Taste of the Tree}
Yeshua said: \\
``The tree is known by its fruit, \\
and the heart is known \\
by the taste of its actions. \\
Do not ask, `Is this path true?' \\
Ask, `What does it ripen in me?' \\
If your seeing bears peace, \\
if your words bear mercy, \\
if your hands bear freedom from suffering, \\
then the root is good, \\
whether you name it Dharma or Kingdom. \\
The wise one does not cling \\
to the name of the tree; \\
he tastes the fruit \\
and understands.''

\ornament

% =========================================================
% CHAPTER 7
% =========================================================
\chapter{Liberation: The Ascent of the Mindful Soul}

\section{Mist Before the Dawn}
``The powers rise \\
like fog along the valley floor— \\
fear, craving, confusion, pride. \\
But the awakened soul says: \\
`You cannot bind me, \\
for I have seen the sun.' \\
At this, \\
the fog remembers \\
what it is \\
and melts. \\
Liberation is not the death of the powers, \\
but their recognition \\
as passing mist.''\\

\section{The Trap of Belief}
``Do not flee the world— \\
the world is a passing cloud. \\
Flee the one \\
who believes the cloud is a prison. \\
For that belief \\
is the iron gate, \\
and clear seeing \\
is the key \\
that has always been in your hand. \\
When you open your fist, \\
you discover \\
there was never a lock.''\\

\section{The Rising Flame}
``When grasping loosens \\
and resistance unwinds, \\
the soul lifts itself \\
as a flame lifts toward the sky— \\
not by effort, \\
but by nature. \\
Do not pull the flame. \\
Do not push the flame. \\
Simply remove the lid \\
from your own heart, \\
and its rising \\
is assured.''

\ornament

% =========================================================
% CHAPTER 8
% =========================================================
\chapter{The Great Return to the Self-That-Is-No-Self}

\section{The Nameless One}
Yeshua said: \\
``When all that you are not \\
falls away like autumn leaves, \\
what remains \\
is the bare tree of your being— \\
rooted in the Nameless One, \\
whose breath is the wind \\
through every branch. \\
Do not cling to the leaves \\
of role and story. \\
Trust the root, \\
and winter itself \\
will feel like home.''\\

\section{Fullness and Emptiness}
``When emptiness is seen clearly, \\
it overflows with fullness. \\
When fullness is held lightly, \\
it reveals its empty heart. \\
Between the two dances \\
the One \\
who is neither and both— \\
like space, \\
bearing all worlds \\
yet never burdened. \\
To awaken is to rest \\
as this silent, shining space.''\\

\section{Already Awakened}
``Blessed is the one \\
who drops even the longing to awaken. \\
For as long as you seek, \\
you imagine distance. \\
At the moment of release, \\
the veil falls, \\
and you see \\
that awakening \\
was the very ground \\
beneath your feet— \\
the Light \\
that never left you, \\
even in your deepest sleep.''

\ornament

\end{document}
