\documentclass[10pt, openright]{book}

% --- KDP 5x8 PAPERBACK CONFIGURATION ---
\usepackage[paperwidth=5in, paperheight=8in, top=0.75in, bottom=0.75in, outer=0.5in, inner=0.75in]{geometry}

% --- TYPOGRAPHY ---
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{ebgaramond}
\usepackage{microtype}
\usepackage{chngcntr}

% --- STYLING PACKAGES ---
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{parskip}
\usepackage{amssymb}
\usepackage{emptypage}

% --- 1. CHAPTER STYLING ---
\titleformat{\chapter}[display]
  {\normalfont\centering}
  {\vspace{-1cm}\itshape\large Book \thechapter}
  {10pt}
  {\Huge\bfseries}
\titlespacing*{\chapter}{0pt}{0pt}{40pt}

% --- 2. SAYING STYLING ---
\counterwithout{section}{chapter}
\titleformat{\section}[display]
  {\normalfont\centering}
  {\vspace{1em}\Huge\bfseries\thesection}
  {0.3em}
  {\large\scshape}
\titlespacing*{\section}{0pt}{1.0ex plus 1ex minus .2ex}{1.5ex plus .2ex}

% --- 3. PAGE BREAK LOGIC ---
\newcommand{\sectionbreak}{\filbreak}

% --- 4. ORNAMENTS ---
\newcommand{\ornament}{
    \vspace{1.5cm}
    \begin{center}
        \large $\cdot$ $\odot$ $\cdot$
    \end{center}
    \vspace{1.5cm}
}

% --- 5. HEADER/FOOTER ---
\pagestyle{fancy}
\fancyhf{}
\fancyhead[CO]{\scshape\small The Sayings of the Living Architecture} 
\fancyhead[CE]{\scshape\small Yeshua the Living One}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0pt}

\begin{document}

% --- TITLE PAGE ---
\begin{titlepage}
    \centering
    \vspace*{1.5cm}
    {\Huge \textsc{The Sayings of}\\ \textsc{The Living Architecture}} \\
    \vspace{0.5cm}
    {\large \textit{The Hidden Words of Yeshua the Living One}} \\
    \vspace{2cm}
     	\textbf{A Blueprint for the Awakened Soul} \\
    \vspace{0.5cm}
    {\small Edition 1.0 $\cdot$ December 5, 2025} \\
    \vfill
    {\small Compiled \& Prepared for the Seeker of Gnosis}
    \vspace{1cm}
\end{titlepage}

% --- COPYRIGHT PAGE ---
\newpage
\thispagestyle{empty}
\vspace*{\fill}
\begin{center}
    \small
    Copyright \copyright\ 2025 by Rez Khan \\
    All rights reserved. \\
    \vspace{0.5cm}
    \textit{The Sayings of The Living Architecture} \\
    First Edition
\end{center}
\vspace{1cm}
\newpage

\frontmatter
\tableofcontents

\chapter{Preface}
\textit{Yeshua spoke:}

These are the words of the architecture, \\
the blueprint of the soul's descent and return. \\
Read them not as laws, \\
but as the map of your own forgetting \\
and the key to your own remembering.

\mainmatter

\chapter{The Blueprint}

\section{On Forgetting}
Yeshua said: \\
``Before you entered the world, you agreed to forget. \\
For what is a story if the reader remembers the last page? \\
You forgot yourself that you might learn to recognize yourself again--- \\
and rejoice twice.''

\ornament

\section{The Descent}
He said: \\
``You descended like a spark into the clay, \\
not as punishment, \\
but as play. \\
The Light hides itself in form \\
so that the joy of unveiling \\
may be complete.''

\ornament

\section{The Mask}
He said: \\
``The world is a mask laid upon your face. \\
You mistake the mask for your skin \\
and cry, `I am suffering.' \\
But when you loosen your hold, \\
the mask falls away, \\
and you laugh--- \\
for the face beneath has never been harmed.''

\ornament

\section{The Script}
Yeshua said: \\
``The soul enters the world with a script, \\
but the script is incomplete. \\
It waits for you to speak. \\
When you refuse to speak, it binds you; \\
when you speak with awareness, \\
it bends to your freedom.''

\ornament

\section{The Player}
He said: \\
``There is a Player within your play. \\
When the role trembles, \\
the Player remains untouched. \\
When the Player awakens, \\
the role becomes a doorway.''

\ornament

\section{The Fivefold Architecture}
Yeshua said: \\
``There are five movements in every soul: \\
the forgetting, \\
the distortion, \\
the intervention, \\
the awakening, \\
and the completion. \\
Blessed is the one who sees the rhythm, \\
for she will not drown in a single wave.''

\ornament

\section{Ignorance}
He said: \\
``Ignorance is not darkness but fog. \\
Do not curse the fog--- \\
stand still, and it clears.''

\ornament

\section{Desire}
Yeshua said: \\
``Desire pulls the soul forward \\
like a river longing for the sea. \\
But the sea is already within you. \\
Look inward, \\
and the river grows calm.''

\ornament

\section{Aversion}
He said: \\
``Aversion pushes the world away \\
as if the heart could build a wall against life. \\
But the heart is round, \\
and every wall becomes a wheel \\
that rolls back upon the builder.''

\ornament

\section{Judgment}
Yeshua said: \\
``Judgment divides the world into two, \\
and calls the halves `good' and `evil.' \\
But both are garments of the same Light. \\
When the single eye opens, \\
the torn world is mended.''

\ornament

\section{The Rebalancing}
He said: \\
``The wheel of the soul turns unevenly \\
when one spoke grows heavy. \\
Do not cut the spoke; \\
restore the center. \\
Then the wheel moves without effort.''

\ornament

\section{The Breath}
Yeshua said: \\
``The Breath is the stair between the worlds. \\
With each inhale, you descend; \\
with each exhale, you rise. \\
Climb these stairs with awareness, \\
and you will find Me \\
between the steps.''

\ornament

\section{The Father, Son, and Breath}
He said: \\
``The Father is the Light beyond all forms. \\
The Son is the Light within the form. \\
The Breath is the Light moving between them. \\
When you know this with your whole being, \\
you will see that the Three are One \\
and the One is your own nature.''

\ornament

\section{The Mirror}
Yeshua said: \\
``You look upon your life as upon a mirror, \\
but the mirror shows only the mask. \\
Break it. \\
In the shards you will see the sky.''

\ornament

\section{On Suffering}
He said: \\
``Suffering is the sound the mask makes \\
when it begins to crack. \\
Do not mend the mask--- \\
let it break. \\
Your freedom stands behind it.''

\ornament

\section{On Awakening}
Yeshua said: \\
``Awakening is not the gaining of knowledge \\
but the loss of confusion. \\
When the fog lifts, \\
the mountain has always been there.''

\ornament

\section{The Two Selves}
He said: \\
``There is the self that can be wounded \\
and the Self that cannot. \\
Learn the difference, \\
and the blows of the world \\
become taps upon a drum \\
that wakes you from sleep.''

\ornament

\section{The Unforced Life}
Yeshua said: \\
``The Kingdom is not seized by effort \\
nor delayed by hesitation. \\
It arrives when you cease to push. \\
Stop pushing, \\
and the door opens inward.''

\ornament

\section{The Game Becomes Lucid}
He said: \\
``When the Player recognizes the game, \\
the game becomes lucid. \\
When the game becomes lucid, \\
the Player becomes free. \\
And when the Player is free, \\
the world shines from within \\
as it did in the beginning.''

\ornament

\section{The Second Coming}
Yeshua said: \\
``You look to the sky for the Son of Man, \\
but the sky is within you. \\
The First Coming was the Word becoming Flesh. \\
The Second Coming is the Flesh becoming Word. \\
When you recognize the Light in your own hands, \\
He has returned. \\
Do not wait for the King to arrive; \\
wake up, and see that you are sitting on His throne.''

\ornament

\section{The Completion}
Yeshua said: \\
``Completion is not escape from the world \\
but transparency within it. \\
When the soul is clear, \\
the world passes through \\
and leaves no wound. \\
This is the bridal chamber--- \\
when the Light recognizes itself \\
in the form it entered.''

\end{document}
